Machine-learned embedded atom method

### 2019-10-11 doc/papers/ml-eam/manuscript.tex
The ML-EAM manuscript is updated.

In this work, we successfully integrated two constraints into the total loss:
the Rose equation of state constraint and the elastic tensor constraint. These 
two losses will be discussed later. The details of their implementations will be 
described in another paper.

The Rose constraint incoporates the universal equation of state (Rose et al) 
into the total loss function. The Mishin-modified equation 
\begin{equation}
\label{eq:rose_eos}
E(x) = E_{0}\left[
    1 + \alpha x + \beta \alpha^3 x^3 \frac{2x + 3}{(x - 1)^2} \right]
    e^{-\alpha x}
\end{equation}
is used because the original form tends to underestimate energies under high 
pressures. In Equation \ref{eq:rose_eos}, $E_{0}$ is the energy of the 
equilibrium structure, $x = a / a_{0} - 1$ is the relative isotropic scaling 
factor ($a$ is the lattice constant), $\beta$ is a chosen parameter and 
\begin{equation}
\label{eq:rose_alpha}
\alpha = \sqrt{-\frac{9 V_{0} B }{E_{0}}}
\end{equation}
where $V_0$ is the equilibrium volume and $B$ is the bulk modulus. The adoption
of the Rose constraint guarantees the exact predictions of bulk modulus and 
the energy-volume curve. The loss of the Rose EOS constraint 
$\mathbf{L}^{\mathrm{Rose}}$  is measured as the 2-norm of the energy 
differences between $E(x)^{\mathrm{Rose}}$ and their corresponding predicted 
$E(x)$: 
\begin{equation}
\label{eq:rose_loss}
\mathbf{L}^{\mathrm{Rose}} =
\sum_{i}{\mathbf{RMSE}(E_{i}(x), E_{i}(x)^{\mathrm{Rose}})}
\end{equation}
In this work, for each included crystal, we use the same choices of $x$: 
$x_{t} = x_{0} + \Delta x \cdot t, x_{0} = -0.1, N_{t}^{\mathrm{max}} = 20, 
\Delta x=0.01$. $\beta$ is fixed to 0.005.

@@

In this work, the loss $\mathbf{L}^{\mathrm{elastic}}$ contributed by the 
elastic tensor is also measured by the RMSE between $c_{ijkl}$ and 
$c_{ijkl}^{\mathrm{dft}}$:
\begin{align}
\label{eq:cijkl_loss}
\mathbf{L}^{\mathrm{elastic}} 
& = \sum_{i}{c_{i} \cdot \mathbf{RMSE_{i}} + {|\epsilon|} + {|\mathbf{F}|}} \\
\label{eq:cijkl_loss_gate}
c_{i} & = \mathbf{ReLU}(\mathbf{MAE_{i}} - \tau) \\
\label{eq:relu}
\mathbf{ReLU}(x) & = \begin{cases}
    x & x \ge 0 \\
    0 & x < 0
\end{cases}
\end{align}
where $\sum_{i}$ loops through all included crystals, $\mathbf{MAE}$ is the mean 
absolute error and $\tau$ is a pre-selected gate parameter. When $\mathbf{MAE}$ 
is below $\tau$, $\mathbf{L}^{\mathrm{elastic}}$ will not contribute to the 
total loss. In this work, $\tau$ is set to 2 GPa.

@@

To find optimal parameters of the funcitons in section \ref{sec:functions}, the 
overall loss function shall be defined. Just like other machine learning tasks, 
the mini-batch stochastic gradient descent algorithm is used to minimize the 
loss function. Equation \ref{eq:loss} demonstrates the loss function used in 
this work:
\begin{align}
\label{eq:loss}
\mathbf{Loss} & = \sqrt{\frac{1}{N_{b}}\sum_{i=1}^{N_{b}}{\left(
    E_{i} - E_{i}^{\mathrm{dft}}
\right)^2}} \nonumber \\
& + \chi_{\mathrm{f}}\sqrt{
    \frac{1}{3\sum_{i}^{N_{b}}{N_i}}\sum_{i}^{N_b}{\sum_{j}^{N_i}{
        \sum_{\alpha}{
            \left(f_{ij\alpha} - f_{ij\alpha}^{\mathrm{dft}}\right)^2
        }
    }}
} \nonumber \\
& + \chi_{\mathrm{s}}\sqrt{\frac{1}{6N_b}\sum_{i}^{N_b}{
    \sum_{j}^{6}{
        \left(
            \epsilon^{\mathrm{voigt}}_{j} - \epsilon^{\mathrm{voigt,dft}}_{j}
        \right)^2
    }
}} \nonumber \\
& + \mathbf{L}^{\mathrm{Rose}} + \mathbf{L}^{\mathrm{elastic}}
\end{align} 
where $N_{b}$ is the batch size and $N_i$ is the number of atoms in structure 
$i$. $\chi_{\mathrm{f}}$ and $\chi_{\mathrm{s}}$ are overall weights of force 
and virial stress contributions. The superscript 'voigt' means that virial 
stress tensors should be converted to Voigt form. The Adam optimizer is used to 
minimize Equation \ref{eq:loss}. In most cases, we use 0.01 as the initial 
learning rate and the batch size ranges from 20 to 50. Since there are very few
adjustable parameters, the optimization typically needs very few epochs to 
converge. When the optimization is finished, the corresponding Lammps setfl 
potential file will be exported.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3. Results
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Results and Discussions}
\label{sec:results}

The publicly available Ni-Mo dataset is use to demonstrate our ML-EAM approach. 
This dataset is built by Shyue Ping Ong and co-workers. It contains 3973 unique
Ni-Mo solids, including 373 elemental Ni and 284 elemental Mo. DFT calculations 
were done by VASP at the PBE level with projector augmented-wave approach.

Table \ref{table:parameters} shows the optimization results. All these 
minimization tasks include energy, force and stress terms in their total losses.
The $\chi_{\mathrm{f}}$ in Equation \ref{eq:loss} is set to 10 and 
$\chi_{\mathrm{s}}$ is fixe to 80. In this paper, the superscript tag 'old' 
denotes the original Zjw04, 'res' means restricted optimization and 'unres' 
means unrestricted optimization. Here 'unrestricted' indicates the parameter 
$r_e$ will be treated as a common adjustable variable.

Table \ref{table:MAE} demonstrates the prediction errors of the machine learned
EAM models and reference models.

\begin{table*}[h]
\centering
\begin{tabular}{lcccccccc}
\hline
    & Ni$^{\mathrm{old}}$ & Ni & Ni$^{\mathrm{alloy}}$ 
    & Mo$^{\mathrm{old}}$ & Mo & Mo$^{\mathrm{alloy}}$ \\
\hline
$r_e$ & 2.488746 & 2.124480 & & 2.728100 & & \\
$f_e$ & 2.007018 & 2.633256 & & 2.723710 & & \\
$\rho_e$ & 27.562015 & 27.233315 & \\ 
$\rho_s$ & 27.930410 & 26.392414 &\\
$\alpha$ & 8.383453 & 8.452753 & \\
$\beta$ & 4.471175 & 3.285651 & \\
$A$ & 0.429046 & 0.9802988 & \\
$B$ & 0.633531 & 0.8919016 & \\
$\kappa$ & 0.443599 & 0.5685785 & \\
$\lambda$ & 0.820658 & 1.1653832 & \\
$F_{n0}$ & -2.693513 & -3.4354472 & \\
$F_{n1}$ & -0.076445 & 0.3544341 & \\
$F_{n2}$ & 0.241442 & -2.5563858 & \\
$F_{n3}$ & -2.375626 & -7.1984844 & \\
$F_{0}$ & -2.70 & -3.236908 & \\
$F_{1}$ & 0 & 1.4576268 & \\
$F_{2}$ & 0.265390 & 2.1785288 & \\
$F_{3}$ & -0.152856 & -1.642411 & \\
$\eta$ & 0.469000 & 4.305329 & \\
$F_e$ & -2.699486 & -3.6163342 & \\
\hline
\end{tabular}
\caption{\label{table:parameters} The original (labeled as 'old'), elemental and 
alloy parameters.}
\end{table*}

>>>>

In this work, we successfully integrated two constraints into the total loss:
the Rose equation of state (EOS) constraint and the elastic tensor constraint. 
These two losses will be discussed later. The details of their implementations 
will be described in another paper.

The Rose constraint incoporates the universal equation of state (Rose et al) 
into the total loss function. The Mishin-modified equation 
\begin{equation}
\label{eq:rose_eos}
E(x) = E_{0}\left[
    1 + \alpha x + \beta \alpha^3 x^3 \frac{2x + 3}{(x - 1)^2} \right]
    e^{-\alpha x}
\end{equation}
is used because the original form tends to underestimate energies under high 
pressures. In Equation \ref{eq:rose_eos}, $E_{0}$ is the energy of the 
equilibrium structure, $x = a / a_{0} - 1$ is the relative isotropic scaling 
factor ($a$ is the lattice constant), $\beta$ is a chosen parameter and 
\begin{equation}
\label{eq:rose_alpha}
\alpha = \sqrt{-\frac{9 V_{0} B }{E_{0}}}
\end{equation}
where $V_0$ is the equilibrium volume and $B$ is the bulk modulus. The adoption
of the Rose constraint guarantees the exact predictions of bulk modulus and 
the energy-volume curve. The loss of the Rose EOS constraint 
$\mathbf{L}^{\mathrm{Rose}}$  is measured as the 2-norm of the energy 
differences between $E(x)^{\mathrm{dft}}$ and their corresponding predicted 
$E(x)$: 
\begin{equation}
\label{eq:rose_loss}
\mathbf{L}^{\mathrm{Rose}} = \sum_{\Omega}{
    \sqrt{\sum_{i}{\left(E(x_i) - E(x_i)^{\mathrm{dft}}\right)^2}}
}
\end{equation}
Here $\Omega$ denotes all selected crystals. In this work, for each included 
crystal, we use the same choices of $x$: 
$x_{i} = x_{0} + \Delta x \cdot i, x_{0} = -0.1, N_{t}^{\mathrm{max}} = 20, 
\Delta x=0.01$. In this work, $\beta^{\mathrm{Ni}}$ is 0.005 and 
$\beta^{\mathrm{Mo}}$ is 0.150.

@@

In this work, the loss $\mathbf{L}^{\mathrm{elastic}}$ contributed by the 
elastic tensor is also measured by the RMSE between $c_{ijkl}$ and 
$c_{ijkl}^{\mathrm{dft}}$:
\begin{align}
\label{eq:cijkl_loss}
\mathbf{L}^{\mathrm{elastic}} 
& = \sum_{\Omega}{\left(
    \omega_{\Omega} \cdot \mathbf{RMSE}_{\Omega} 
    + ||\epsilon|| + ||\mathbf{F}||
\right)}
 \\
\label{eq:cijkl_loss_gate}
\omega_{\Omega} & = \mathbf{ReLU}(\mathbf{MAE}_{\Omega} - \tau) \\
\label{eq:relu}
\mathbf{ReLU}(x) & = \begin{cases}
    x & x \ge 0 \\
    0 & x < 0
\end{cases}
\end{align}
where $\sum_{\Omega}$ loops through all included crystals, 
$\mathbf{MAE}_{\Omega}$ and $\mathbf{RMSE}_{\Omega}$ are the mean absolute error 
and root mean squared error between predicted elastic constants and DFT elastic 
constants. $\tau$ is a pre-selected gate parameter. 
When $\mathbf{MAE}$ is below $\tau$, $\mathbf{L}^{\mathrm{elastic}}$ will not 
contribute to the total loss. In this work, $\tau$ is set to 2 GPa.

@@

To find optimal parameters of the funcitons in section \ref{sec:functions}, the 
overall loss function shall be defined. Just like other machine learning tasks, 
the mini-batch stochastic gradient descent algorithm is used to minimize the 
loss function. Equation \ref{eq:loss} demonstrates the loss function used in 
this work:
\begin{align}
\label{eq:loss}
\mathbf{Loss} & = \sqrt{\frac{1}{N_{b}}\sum_{i=1}^{N_{b}}{\left(
    E_{i} - E_{i}^{\mathrm{dft}}
\right)^2}} \nonumber \\
& + \chi_{\mathrm{f}}\sqrt{
    \frac{1}{3\sum_{i}^{N_{b}}{N_i}}\sum_{i}^{N_b}{\sum_{j}^{N_i}{
        \sum_{\alpha}{
            \left(f_{ij\alpha} - f_{ij\alpha}^{\mathrm{dft}}\right)^2
        }
    }}
} \nonumber \\
& + \chi_{\mathrm{s}}\sqrt{\frac{1}{6N_b}\sum_{i}^{N_b}{
    \sum_{j}^{6}{
        \left(
            \epsilon^{\mathrm{voigt}}_{j} - \epsilon^{\mathrm{voigt,dft}}_{j}
        \right)^2
    }
}} \nonumber \\
& + \chi_{\mathrm{rose}}\mathbf{L}^{\mathrm{Rose}} 
+ \chi_{\mathrm{elastic}}\mathbf{L}^{\mathrm{elastic}}
\end{align} 
where $N_{b}$ is the batch size and $N_i$ is the number of atoms in structure 
$i$. $\chi_{\mathrm{f}}$, $\chi_{\mathrm{s}}$, $\chi_{\mathrm{rose}}$ and 
$\chi_{\mathrm{elastic}}$ are overall weights of force, virial stress, rose EOS 
and elastic contributions. The superscript 'voigt' means that virial 
stress tensors should be converted to Voigt form. The Adam optimizer is used to 
minimize Equation \ref{eq:loss}. In most cases, we use 0.01 as the initial 
learning rate and the batch size ranges from 20 to 50. Since there are very few
adjustable parameters, the optimization typically needs very few epochs to 
converge. When the optimization is finished, the corresponding Lammps setfl 
potential file will be exported.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3. Results
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Results}
\label{sec:results}

The publicly Ni-Mo dataset is used to evaluate our approach. This dataset is 
provided by Shyue Ping Ong. It contains 3973 different Ni-Mo solids. All 
calculations were done by VASP with the PBE functional and the projector 
augmented-wave approach.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.A
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Ni}
\label{sec:elementary_Ni}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Ni - EAM - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure*}[h!]
\centering
\includegraphics[scale=0.57]{figures/Ni_eam.png}
\caption{\label{fig:Ni_eam} $\rho(r)$, $F(\rho)$ and $\phi(r)$ of the original 
Zjw04 EAM and the machine learned EAM (ML-EAM).}
\end{figure*}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Ni - EOS
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[h!]
\centering
\includegraphics[scale=0.65]{figures/Ni_all_eos.png}
\caption{\label{fig:Ni_eam} $\rho(r)$, $F(\rho)$ and $\phi(r)$ of the original 
Zjw04 EAM and the machine learned EAM (ML-EAM).}
\end{figure}

We first use the elementary fcc Ni dataset to show our approach. This dataset 
has 461 structures in total. 400 were randomly selected to form the training 
subset. In this experiment, $\chi_{\mathrm{f}}$, $\chi_{\mathrm{rose}}$ and 
$\chi_{\mathrm{elastic}}$ are set to 1, 3 and 0.05 respectively. In this 
section, SNAP shall refer to the elementary SNAP Ni model. Table 
\ref{table:elementary_Ni_mae} shows the MAEs of those models. The original EAM 
model already has an accepetable performace on this dataset. Our machine 
learning approach can further reduce the energy MAE to 4.1 meV/atom. All of 
these models have similar force prediction performance.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table: Ni - MAE
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table}[h]
\centering
\begin{tabular}{lccc}
\hline
    & SNAP & EAM & ML-EAM \\
\hline
Energy MAE (meV/atom) & 1.2 & 10.6 & 4.1 \\
Force MAE (eV/\AA) & 0.05 & 0.06 & 0.05 \\
\hline
\end{tabular}
\caption{\label{table:elementary_Ni_mae} Comparion of the energy and force 
accuracy of the elementary SNAP Ni mode, the original EAM and the machine 
learned EAM.
}
\end{table}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table: Ni - Material Properties
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table*}[h]
\centering
\begin{tabular}{lcccccc}
\hline
    & DFT & SNAP & EAM & ML-EAM & Expt. \\
\hline
$T_{m}$(K) & & 1785 & 1520 & 1765 & 1600 & 1728 \\
$c_{11}$ (GPa) & 276 & 276 (0.0\%) & 248 (-10.1\%) & 260 (-5.8\%) & 274 (-0.7\%) & 261 \\
$c_{12}$ (GPa) & 159 & 159 (0.0\%) & 147 (-7.5\%) & 151 (0.0\%) & 163 (2.5\%) & 151 \\
$c_{44}$ (GPa) & 132 & 132 (0.0\%) & 125 (-5.3\%) & 131 (0.0\%) & 131 (-0.8\%) & 132 \\
$E_{v}$ (eV) & 1.46 & 1.68 (15.1\%) & 1.68 (15.1\%) & 1.16 (-20.5\%) & 1.71 (17.1\%) & 1.54–1.80 \\
$E_{m}$ (eV) & 1.12 & 1.07 (-4.5\%) & 0.90 (-19.6\%) & 1.46 (30.5\%) & 0.87 (-22.3\%) & 1.01–1.48 \\
$E_{a} = E_{v} + E_{m} $ (eV) & 2.58 & 2.75 (6.6\%) & 2.58 (0.0\%) & 2.62 (1.6\%) & 2.58 (0.0\%) & 2.77–2.95 \\
\hline
\end{tabular}
\caption{\label{table:elementary_Ni} Comparison of the calculated and 
experimental material properties of elementary Ni. $T_{m}$ is the melting 
point. $c_{ij}$ represents elastic constants. $E_{v}$ and $E_{m}$ are vacancy 
formation energy and migration energy, respectively. SNAP refers to the 
elementary Ni SNAP model.
}
\end{table*}

% Table \ref{table:parameters} shows the optimization results. All these 
% minimization tasks include energy, force and stress terms in their total losses.
% The $\chi_{\mathrm{f}}$ in Equation \ref{eq:loss} is set to 10 and 
% $\chi_{\mathrm{s}}$ is fixe to 80. In this paper, the superscript tag 'old' 
% denotes the original Zjw04, 'res' means restricted optimization and 'unres' 
% means unrestricted optimization. Here 'unrestricted' indicates the parameter 
% $r_e$ will be treated as a common adjustable variable.

% Table \ref{table:MAE} demonstrates the prediction errors of the machine learned
% EAM models and reference models.

% \begin{table*}[h]
% \centering
% \begin{tabular}{lcccccccc}
% \hline
%     & Ni$^{\mathrm{old}}$ & Ni & Ni$^{\mathrm{alloy}}$ 
%     & Mo$^{\mathrm{old}}$ & Mo & Mo$^{\mathrm{alloy}}$ \\
% \hline
% $r_e$ & 2.488746 & 2.124480 & & 2.728100 & & \\
% $f_e$ & 2.007018 & 2.633256 & & 2.723710 & & \\
% $\rho_e$ & 27.562015 & 27.233315 & \\ 
% $\rho_s$ & 27.930410 & 26.392414 &\\
% $\alpha$ & 8.383453 & 8.452753 & \\
% $\beta$ & 4.471175 & 3.285651 & \\
% $A$ & 0.429046 & 0.9802988 & \\
% $B$ & 0.633531 & 0.8919016 & \\
% $\kappa$ & 0.443599 & 0.5685785 & \\
% $\lambda$ & 0.820658 & 1.1653832 & \\
% $F_{n0}$ & -2.693513 & -3.4354472 & \\
% $F_{n1}$ & -0.076445 & 0.3544341 & \\
% $F_{n2}$ & 0.241442 & -2.5563858 & \\
% $F_{n3}$ & -2.375626 & -7.1984844 & \\
% $F_{0}$ & -2.70 & -3.236908 & \\
% $F_{1}$ & 0 & 1.4576268 & \\
% $F_{2}$ & 0.265390 & 2.1785288 & \\
% $F_{3}$ & -0.152856 & -1.642411 & \\
% $\eta$ & 0.469000 & 4.305329 & \\
% $F_e$ & -2.699486 & -3.6163342 & \\
% \hline
% \end{tabular}
% \caption{\label{table:parameters} The original (labeled as 'old'), elemental and 
% alloy parameters.}
% \end{table*}
